\vspace{-0.8in}Brief Announcement: A Family of Leaderless Generalized-Consensus Algorithms

### 2016-02-12 podc16/paper2.tex
formatting changes

We now present our abstract leaderless generalized consensus algorithm.
A process $p$ receiving a command $c$ proceeds in three phases. First, it calls a \textit{dependency-set} sub-algorithm (\Cref{dep-algo}) to determine a set of commands $\mathcal{D}$ that the command $c$ may depend on. Second, $p$ proposes the mapping  $c\mapsto \mathcal{D}$ to a \textit{map-agreement} sub-algorithm (\Cref{map-algo}); the map-agreement sub-algorithm may subsequently commit a mapping $c\mapsto \mathcal{D}'$ for $c$, in which case $\mathcal{D}'$ becomes the dependency set of $c$;
we say that $c$ is executable when all its transitive dependencies are known. %no command $c'$ for which no mapping has been committed is reachable from $c$ by recursively following committed mappings. 
\todo{Change the name of the the decide action to commit.}
Finally, when $c$ is executable, the $p$ may execute $c$ by deciding the c-struct obtained by \emph{locally} running a \emph{graph-processing} algorithm described in \Cref{abstract-algo}.
Any other process can run the three phases independently of $p$ and eventually execute the command despite a crash of $p$.
However, we show that a leaderless  generalized consensus algorithm is always safe and is live when its underlying dependency-set and map-agreement algorithms are safe and live.
The leaderless generalized consensus algorithm orchestrates the two dependency-set and map-agreement sub-algorithms as shown in \cref{fig:deps-set-algo3}.\todo{This pseudocode must be in figure (a). Also we have to better describe the pseudocode.}

@@

A dependency-set algorithm exposes the input interface $\opn{announce}_p\left( c \right)$, to announce a command $c$, the output interface $\opn{suggest}_p \left( c,\mathcal{D}\right)$, to suggest a set of dependencies $\mathcal{D}$ for $c$, and \todo{this is not clear: an input interface $\opn{commit}\left( c,\mathcal{D} \right)$ (meant to receive inputs from the agreement sub-algorithm)}, where $p\in P$ is the process executing the call.
Moreover, a dependency-set algorithm must ensure that:
\begin{enumerate}[noitemsep,nolistsep]
    \item For any suggestion $\opn{suggest}_p \left( c,\mathcal{D}\right)$, every command in $\{c\}\cup \mathcal{D}$ must have been announced before.
    \item For any two suggestions $\opn{suggest}_p \left( c_1,\mathcal{D}_1\right)$ and $\opn{suggest}_q \left( c_2,\mathcal{D}_2\right)$ (with $p$ possibly equal to $q$), if $c_1$ and $c_2$ do not commute then either $c_1\in \mathcal{D}_2$ or $c_2 \in \mathcal{D}_1$.
    \item If a suggestion $\opn{suggest}_p(c,\mathcal{D})$ has been made and no input $\opn{commit}_q\left( c,\mathcal{D'} \right)$ is observed, then eventually a new suggestion $\opn{suggest}_r\left( c,\mathcal{D''} \right)$ for $r\neq p$ is made.
\end{enumerate}

\begin{figure*}
\centering
\subfigure[$\texttt{init}_p$ and $\texttt{announce}_p$ implementations.]{
\noindent\fbox{%
	\begin{varwidth}{\dimexpr\linewidth-2\fboxsep-2\fboxrule\relax}

@@

		\begin{algorithmic}[1]
			\makeatletter
			\setcounter{ALG@line}{8}
			\makeatother
			{\scriptsize
			\START[$\langle \texttt{cmd}$ $|$ $c \rangle$]{{\bf upon receive}}{ {\bf from} $q$} \label{alg:receive-new-cmd}
			\State $\mathcal{D} \gets \{ d : d \in cmds_p \land$\label{alg:update-d}
			\Statex \hspace{2cm}$\land d\neq c \land d \asymp c \}$
			\State $cmds_p\gets cmds_p\cup\{c\}$\label{alg:update-cmds-p}
			\State {\bf send} $\langle \texttt{deps}$ $|$ $c,\mathcal{D} \rangle$ {\bf to} $q$\label{alg:send-deps}
			\END
			\Statex
			\START[$\langle \texttt{deps}$ $|$ $c,\mathcal{D} \rangle$]{{\bf upon receive}}{ {\bf from} $q$} \label{alg:receive-deps}
			\State $deps_p[c]\gets deps_p[c]\cup \mathcal{D}$\label{alg:update-deps-p}
			\State $heard_p[c]\gets heard_p[c]\cup \{q\}$
			\END
			\Statex
			\START[]{{\bf upon}}{$\exists c$ $:$ $ | heard_p[c] | \geq \mathcal{Q}$ $\land$} \label{alg:receive-deps}
			\Statex \hspace{1cm}$\land$ $suggested_p[c]=false$
			\State $suggested_p[c]\gets true$
			\State {\bf call} $\texttt{suggest}_p(c,deps_p[c])$ \label{alg:call-suggest}
			\END
			}
		\end{algorithmic}
	\end{varwidth}%
}
\label{fig:deps-set-algo2}
}
\hspace*{0.1in}\subfigure[Leaderless generalized consensus implementation.]{
\noindent\fbox{%
	\begin{varwidth}{\dimexpr\linewidth-2\fboxsep-2\fboxrule\relax}

		\begin{algorithmic}[1]
			{\scriptsize
			\START[$\opn{init}_p$]{{\bf upon}}{$()$}
			\State $m_p\gets \emptyset$
			\State $\sigma_p\gets \bot$
			\END
			\START[$\opn{GC-propose}_p$]{{\bf upon}}{$(c)$}
			\State {\bf call} $\opn{announce}_p\left( c \right)$
			\END
			\Statex
			\START[$\opn{suggest}_p$]{{\bf upon}}{$(c,\mathcal{D})$}
			\State  $\opn{propose}_p( c,\mathcal{D})$ 			
			\END
			\Statex
			\START[$\opn{commit}_p$]{{\bf upon}}{$(c,\mathcal{D})$}
			\State  $m_p \gets m_p \oplus \left[ c \mapsto \mathcal{D}\right]$	
			\State $\sigma_p \gets  \opn{c-struct}\left(m_p\right)$
			\State $\opn{GC-decide}_p(\sigma_p)$		
			\END
			}
		\end{algorithmic}
	\end{varwidth}%
}
\label{fig:deps-set-algo3}
}
\caption{Example of dependency-set algorithm and leaderless generalized consensus implementations.}
\label{fig:deps-set-algo}
\end{figure*}

@@

A map-agreement algorithm exposes the input interface $\opn{propose}_p\left( c, \mathcal{D} \right)$, to propose a set of dependencies $\mathcal{D}$ for a command $c$, and the output interface $\opn{commit}_p\left( c, \mathcal{D} \right)$, to commit $\mathcal{D}$ for $c$, where $p\in P$ is the process executing the call.
Moreover, a map-agreement algorithm must ensure that: 
\begin{comment}
\begin{enumerate}[noitemsep,nolistsep]
    \item If the set of dependencies $\mathcal{D}$ is committed for a command $c$, then $\mathcal{D}$ has been proposed for $c$ at an earlier time.
    \item If $\mathcal{D}_1$ is committed for a command $c$ and $\mathcal{D}_2$ is committed for the same command $c$, then $\mathcal{D}_1$ is equal to $\mathcal{D}_2$.
 \end{enumerate}

\todo{OR MORE FORMALLY}

@@

\todo{Cite Generalized Paxos as an example for bounding the data structures. Problem: we do not describe any particular data structure here. It is not clear why we should write about GC here.}

@@

>>>>

Our leaderless generalized consensus algorithm orchestrates a dependency-set sub-algorithm and map-agreement sub-algorithm as shown in \cref{fig:deps-set-algo3}.
A process $p$ receiving a command $c$ proceeds in three phases. First, it calls a \textit{dependency-set} sub-algorithm (\Cref{dep-algo}) to determine a set of commands $\mathcal{D}$ that the command $c$ may depend on. Second, $p$ proposes the mapping  $c\mapsto \mathcal{D}$ to a \textit{map-agreement} sub-algorithm (\Cref{map-algo}); the map-agreement sub-algorithm may subsequently commit a mapping $c\mapsto \mathcal{D}'$ for $c$, in which case $\mathcal{D}'$ becomes the dependency set of $c$;
we say that $c$ is executable when all its transitive dependencies are known. %no command $c'$ for which no mapping has been committed is reachable from $c$ by recursively following committed mappings. 
Finally, when $c$ is executable, the $p$ may execute $c$ by deciding the c-struct obtained by \emph{locally} running a \emph{graph-processing} algorithm described in \Cref{abstract-algo}.
Any other process can run the three phases independently of $p$ and eventually execute the command despite a crash of $p$.
However, we show that a leaderless  generalized consensus algorithm is always safe and is live when its underlying dependency-set and map-agreement algorithms are safe and live.

@@

A dependency-set algorithm exposes the input interface $\opn{announce}_p\left( c \right)$, to announce a command $c$, the output interface $\opn{suggest}_p \left( c,\mathcal{D}\right)$, to suggest a set of dependencies $\mathcal{D}$ for $c$, and an input interface $\opn{commit}\left( c,\mathcal{D} \right)$ (meant to observe the map-agreement sub-algorithm), where $p\in P$ is the process executing the call.
Moreover, a dependency-set algorithm must ensure that:
\begin{itemize}[noitemsep,nolistsep]
    \item[-] Safety: (a) For any suggestion $\opn{suggest}_p \left( c,\mathcal{D}\right)$, every command in $\{c\}\cup \mathcal{D}$ must have been announced before, and (b), for any two suggestions $\opn{suggest}_p \left( c_1,\mathcal{D}_1\right)$ and $\opn{suggest}_q \left( c_2,\mathcal{D}_2\right)$, if $c_1$ and $c_2$ do not commute then either $c_1\in \mathcal{D}_2$ or $c_2 \in \mathcal{D}_1$.
%    \item If $c$ was announced by $p$, then eventually a dependency set is suggested for $c$ to $p$.
    \item[-] Liveness: (a) If a suggestion $\opn{suggest}_p\left( c \right)$ was made, then eventually a suggestion $\opn{suggest}\left( c,\mathcal{D} \right)_p$ is made, and (b), if a suggestion $\opn{suggest}_p(c,\mathcal{D})$ has been made and no input $\opn{commit}_q\left( c,\mathcal{D'} \right)$ is observed, then eventually a new suggestion $\opn{suggest}_r\left( c,\mathcal{D''} \right)$ for $r\neq p$ is made.
\end{itemize}
\vspace{5cm}

\begin{figure*}
\centering
\subfigure[Leaderless generalized consensus implementation.]{
\noindent\fbox{%
	\begin{varwidth}{\dimexpr\linewidth-2\fboxsep-2\fboxrule\relax}

		\begin{algorithmic}[1]
			{\scriptsize
			\START[$\opn{init}_p$]{{\bf upon}}{$()$}
			\State $m_p\gets \emptyset$
			\State $\sigma_p\gets \bot$
			\END
			\START[$\opn{GC-propose}_p$]{{\bf upon}}{$(c)$}
			\State {\bf call} $\opn{announce}_p\left( c \right)$
			\END
			\Statex
			\START[$\opn{suggest}_p$]{{\bf upon}}{$(c,\mathcal{D})$}
			\State  $\opn{propose}_p( c,\mathcal{D})$ 			
			\END
			\Statex
			\START[$\opn{commit}_p$]{{\bf upon}}{$(c,\mathcal{D})$}
			\State  $m_p \gets m_p \oplus \left[ c \mapsto \mathcal{D}\right]$	
			\State $\sigma_p \gets  \opn{c-struct}\left(m_p\right)$
			\State $\opn{GC-decide}_p(\sigma_p)$		
			\END
			}
		\end{algorithmic}
	\end{varwidth}%
}
\label{fig:deps-set-algo3}
}
\subfigure[$\texttt{init}_p$ and $\texttt{announce}_p$ implementations.]{
\noindent\fbox{%
	\begin{varwidth}{\dimexpr\linewidth-2\fboxsep-2\fboxrule\relax}

@@

		\begin{algorithmic}[1]
			\makeatletter
			\setcounter{ALG@line}{8}
			\makeatother
			{\scriptsize
			\START[$\langle \texttt{cmd}$ $|$ $c \rangle$]{{\bf upon receive}}{ {\bf from} $q$} \label{alg:receive-new-cmd}
			\State $\mathcal{D} \gets \{ d : d \in cmds_p \land$\label{alg:update-d}
			\Statex \hspace{2cm}$\land d\neq c \land d \asymp c \}$
			\State $cmds_p\gets cmds_p\cup\{c\}$\label{alg:update-cmds-p}
			\State {\bf send} $\langle \texttt{deps}$ $|$ $c,\mathcal{D} \rangle$ {\bf to} $q$\label{alg:send-deps}
			\END
			\Statex
			\START[$\langle \texttt{deps}$ $|$ $c,\mathcal{D} \rangle$]{{\bf upon receive}}{ {\bf from} $q$} \label{alg:receive-deps}
			\State $deps_p[c]\gets deps_p[c]\cup \mathcal{D}$\label{alg:update-deps-p}
			\State $heard_p[c]\gets heard_p[c]\cup \{q\}$
			\END
			\Statex
			\START[]{{\bf upon}}{$\exists c$ $:$ $ | heard_p[c] | \geq \mathcal{Q}$ $\land$} \label{alg:receive-deps}
			\Statex \hspace{1cm}$\land$ $suggested_p[c]=false$
			\State $suggested_p[c]\gets true$
			\State {\bf call} $\texttt{suggest}_p(c,deps_p[c])$ \label{alg:call-suggest}
			\END
			}
		\end{algorithmic}
	\end{varwidth}%
}
\label{fig:deps-set-algo2}
}
\caption{Example of dependency-set algorithm and leaderless generalized consensus implementations.}
\label{fig:deps-set-algo}
\end{figure*}

@@

A map-agreement algorithm exposes the input interface $\opn{propose}_p\left( c, \mathcal{D} \right)$, to propose a set of dependencies $\mathcal{D}$ for a command $c$, and the output interface $\opn{commit}_p\left( c, \mathcal{D} \right)$, to commit $\mathcal{D}$ for $c$, where $p\in P$ is the process executing the call.
Moreover, a map-agreement algorithm must ensure that:

@@

@@

\subsection{Description map-agreement implementation}
